Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-02-13 vertical-thermal-review.Rmd
update table 2
#100
update table 1 references
update table 1
#100
added a rought leaf optical property section
Krista, I think your suggestion actually is better. Maybe this would be better as #3.2?

@@

$\ddagger$ 1. Kafuti et al. 2020; 2. Van Wittenberghe et al. 2012; 3. Roberts et al. 1990; 4. Dang et al. 1997; 5. Marenco et al. 2017; 6. Ambrose et al. 2015; 7. Zweifel et al. 2001; 8. Slot et al. 2019; 9. Hernandez et al. 2020; 10. Urban et al. 2007; 11. Carter and Cavaleri 2018; 12. Martin et al. 1999; 13. Mau et al. 2018;  14. Kosugi et al. 2012; 15. Niinemets et al. 2015; 16. Bachofen et al. 2020; 17. Hamerlynck and Knapp 1994; 18. Coble et al. 2017; 19. Wyka et al. 2012; 20. Rijkerse et al. 2000; 21. Ishida et al. 1999; 22. Weerasinghe et al. 2014; 23. Scartazza et al. 2016; 24. Miller et al. 2021; 25. Harris and Medina 2013; 26. Legner et al. 2014; 27. Kitao et al. 2012; 28. Fauset et al. 2018; 29. Rey-Sanchez et al. 2016; 30. Muller et al. 2021; 31. Mier et al. 2001; 32. Turnbull et al. 2003; 33. Araki et al. 2017; 34. Bolstad et al. 1999; 35. Kenzo et al. 2015; 36. Harley et al. 1996; 37. Xu and Griffin 2006; 38. Atherton et al. 2017; 39. Carter et al. 2021; 40. Sack et al. 2003; 41. Slot et al. 2021; 42. Carswell et al. 2000
```{r eval=FALSE, echo=FALSE}
# refs in this table currently in the nocite field at top of document 
```

@@

## 3.3 – Light absorption and reflectance

@@

$\ddagger$ 1. Mau et al. 2018; 2. Coble and Cavaleri 2014; 3. Sack et al. 2006; 4. Chin and Sillett 2019; 5. Wyka et al. 2012; 6. Atherton et al. 2017; 7. Kenzo et al. 2015; 8. Kusi and Karasi 2020; 9. Dang et al. 1997; 10. Gebauer et al. 2015; 11. Marenco et al. 2017; 12. Kafuti et al. 2020; 13. Van Wittenberghe et al. 2012; 14. Zhang et al. 2019; 15. Weerasinghe et al. 2014; 16. Oldham et al. 2010; 17. Ichie et al. 2016; 18. Gregoriou et al. 2007; 19. Levizou et al. 2005; 20. Liakoura 1997; 21. Fauset et al. 2018; 22. Niinemets et al. 1998, 23. Ishida et al. 1998; 24. Millen and Clendon 1979; 25. Smith and Carter, 1988; 26. Hadley and Smith 1987; 28. Baltzer and Thomas 2005; 29. Coble et al. 2016; 30. Scartazza et al. 2016; 31. Duursma and Marshall, 2006; 32. Harley et al. 1996;  33. Hernandez et al. 2020; 34. Turnbull et al. 2003; 35. Chen et al. 2020; 36. van de Weg et al. 2012; 37. M.A Cavaleri et al. 2008;  38. Koniger et al. 1995; 39. Mastubara et al. 2009; 40. Harris and Medina 2013; 41. Hansen et al. 2001; 42. Poorter et al. 1995; 43. Coble et al. 2016; 44. Niinemets et al. 2004; 45. Poorter et al. 2000; 46. Zwieniecki et al. 2004; 47. Sack and Scoffoni, 2013; 48. Ball et al., 1988; 49. Taylor et al. 2021; 50. Niinemets et al. 2010; 51. Brooks et al. 1997; 52. Curtis et al. 2019; 53. Slot et al. 2019; 54. Hamerlynck and Knapp 1994; 55. Lloyd et al. 2010; 56. Augspurger and Bartlett, 2003; 57. Osada et al. 2001; 58. Meakem et al. 2018; 59. Condit et al. 2000; 60. Harley et al. 1997; 61. Niinemets and Sun, 2014; 62. Sharkey and Monson, 2014; 63. Simpraga et al. 2013

>>>>

![](table1c.png)

@@

$\ddagger$ 1. Mau et al. 2018; 2. Coble and Cavaleri 2014; 3. Sack et al. 2006; 4. Chin and Sillett 2019; 5. Wyka et al. 2012; 6. Atherton et al. 2017; 7. Kenzo et al. 2015; 8. Kusi and Karasi 2020; 9. Dang et al. 1997; 10. Gebauer et al. 2015; 11. Marenco et al. 2017; 12. Kafuti et al. 2020; 13. Van Wittenberghe et al. 2012; 14. Zhang et al. 2019; 15. Weerasinghe et al. 2014; 16. Oldham et al. 2010; 17. Ichie et al. 2016; 18. Gregoriou et al. 2007; 19. Levizou et al. 2005; 20. Liakoura 1997; 21. Fauset et al. 2018; 22. Niinemets et al. 1998, 23. Ishida et al. 1998; 24. Millen and Clendon 1979; 25. Smith and Carter, 1988; 26. Hadley and Smith 1987; 28. Baltzer and Thomas 2005; 29. Coble et al. 2016; 30. Scartazza et al. 2016; 31. Duursma and Marshall, 2006; 32. Harley et al. 1996;  33. Hernandez et al. 2020; 34. Turnbull et al. 2003; 35. Chen et al. 2020; 36. van de Weg et al. 2012; 37. M.A Cavaleri et al. 2008;  38. Koniger et al. 1995; 39. Mastubara et al. 2009; 40. Harris and Medina 2013; 41. Hansen et al. 2001; 42. Poorter et al. 1995; 43. Coble et al. 2016; 44. Niinemets et al. 2004; 45. Poorter et al. 2000; 46. Zwieniecki et al. 2004; 47. Sack and Scoffoni, 2013; 48. Ball et al., 1988; 49. Taylor et al. 2021; 50. Niinemets et al. 2010; 51. Brooks et al. 1997; 52. Curtis et al. 2019; 53. Slot et al. 2019; 54. Hamerlynck and Knapp 1994; 55. Lloyd et al. 2010; 56. Augspurger and Bartlett, 2003; 57. Osada et al. 2001; 58. Meakem et al. 2018; 59. Condit et al. 2000; 60. Harley et al. 1997; 61. Niinemets and Sun, 2014; 62. Sharkey and Monson, 2014; 63. Simpraga et al. 2013
```{r eval=FALSE, echo=FALSE}
# refs in this table currently in the nocite field at top of document 
```

@@

## 3.3 – Leaf optical properties
<!---need to cut parts of this paragraph or merge into 3.1?-->
Leaf optical properties of leaves are influenced by a combination of anatomical, morphological, biochemical traits outlined above and water content, that vary vertically.
Leaf light absorptance is maximized along the vertical light gradient and is balanced with leaf temperature regulation. Yet, leaf absorptance is only modestly greater in sun leaves. In the understory or shade regions where PAR is more limiting, leaf light adsorption efficiency is greater than in sun leaves, enhanced by shade leaves maintaining higher SLA, denser layers of chloroplasts, shorter, fewer layers of conical palisade cells, higher nitrogen, and chlorophyll per unit biomass that exemplifies optical absorption capabilites, which realizes to be similar chlorophyll concentration per unit area as in sun leaves. Therefore, shade leaves can have comparable light harvesting capabilities as in sun leaves, and higher absorption effiency per unit biomass [@poorterLeafOpticalProperties1995; @poorterLeafOpticalProperties2000; @leeCorrelatesLeafOptical1990]. On the other hand, in higher light environment, there is less need for light absorption efficiency, but greater need to channel higher availablity of light into deeper cell layers through "light pipes" or palisade layers that increase path length for light penetration, contributing to higher photosynthetic capacity. At the same time, PAR absorptance and thus heat load is reduced by reflecting light through leaf surface modifications as mentioned above, such as trichome and cuticle layer, or biochemically dissipating heat through caretenoid concentrations [Table 1, @knappVariabilityLeafOptical1998; @leeCorrelatesLeafOptical1990]. Thicker sun leaves also reduce light transmittance or light scattered after entering the leaf than shade leaves [@poorterLeafOpticalProperties1995; @leeCorrelatesLeafOptical1990]. Therefore, while canopy leaves tend to invest more energy in metabolic activity and heat dissipation, understory leaves invest in light capture.

@@

$\ddagger$ 1. Kafuti et al. 2020; 2. Van Wittenberghe et al. 2012; 3. Roberts et al. 1990; 4. Dang et al. 1997; 5. Marenco et al. 2017; 6. Ambrose et al. 2015; 7. Zweifel et al. 2001; 8. Slot et al. 2019; 9. Hernandez et al. 2020; 10. Urban et al. 2007; 11. Carter and Cavaleri 2018; 12. Martin et al. 1999; 13. Mau et al. 2018;  14. Kosugi et al. 2012; 15. Niinemets et al. 2015; 16. Bachofen et al. 2020; 17. Hamerlynck and Knapp 1994; 18. Coble et al. 2017; 19. Wyka et al. 2012; 20. Rijkerse et al. 2000; 21. Ishida et al. 1999; 22. Weerasinghe et al. 2014; 23. Scartazza et al. 2016; 24. Miller et al. 2021; 25. Harris and Medina 2013; 26. Legner et al. 2014; 27. Kitao et al. 2012; 28. Fauset et al. 2018; 29. Rey-Sanchez et al. 2016; 30. Muller et al. 2021; 31. Mier et al. 2001; 32. Turnbull et al. 2003; 33. Araki et al. 2017; 34. Bolstad et al. 1999; 35. Kenzo et al. 2015; 36. Harley et al. 1996; 37. Xu and Griffin 2006; 38. Atherton et al. 2017; 39. Carter et al. 2021; 40. Sack et al. 2003; 41. Slot et al. 2021; 42. Carswell et al. 2000
